---
title: "Modern Dive Book Club"
author: "r4ds"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#0B5CA9",
  link_color = "#71BD4C",
  header_font_google = google_font("Helvetica Neue"),
  text_font_google   = google_font("Helvetica Neue", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```


# Agenda

- Introductions
--

- Discuss desired frequency and format for sessions
--

- Book outline
--

- Determine starting point
--

- Presentation sign-ups
--

- Resources

---
class: inverse, center, middle

# Introductions

Name 

Goal 

etc.

---
class: inverse, center, middle

# Frequency


A chapter per week?

---

# Format

2 parts

---

# Part 1 Chapter Presentation 
### (20-30 min, but could last entire time)

--

1. Concepts (with concrete examples)
--

2. Pose questions to the group (confusions)
--

3. Live coding of example/extensions/exercises 
--

4. Conclude with a big picture diagram
    
---

# Part 2 - Questions, extensions, applications
### presenter-led (or facilitator as fallback) 

--

1. Reviewing exercises in book 
--

2. Relating concepts to other parts of the book
--

3. Relating concepts to work you've done before
--

4. Showing something you've built that applies the concepts (e.g, functions, shiny apps, packages)

---
class: inverse, center, middle

# Sound good?

---

# Book outline

1. Data Science w/ tidyverse
--

2. Data Modeling with moderndive 
--

3. Statistical Inference with infer
--

4. Conclusion

---

class: inverse, center, middle

# Determine starting point

Do we want to do the data science + tidyverse intro?

---
class: inverse, center, middle

# Presentation signups

Anyone?

---

# Resources

* [Install R](https://www.r-project.org/)
* [Install R Studio](https://rstudio.com/products/rstudio/download/)
